---
title: "Cupcake V. Muffin EDA"
output: html_notebook
---

Now that I have a dataset of the ingredients of many cupcake and muffin recipes, it's time for phase 2---EDA.

I already know it is a pretty sparse data set and that the number of cupcakes and muffins are not very even.

Figures I might want to make:

- "heatmap" of pres/abs of ingredients.
- boxplot of calories with scatterplot and labeled outliers
- ridge plot of distribution of amounts per serving of each ingredient


**IMPORTANT** Don't use calories with cupcake recipes because you've removed the frosting!
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggridges)
library(iheatmapr)
library(gglabeller)
```

# Read in files
```{r}
# With frosting and calories/serving
recipes <- read_rds("recipes_tidy.rds")
# without frosting and calories
nofrosting <- read_rds("nofrosting_tidy.rds")
```

# Number of Recipes
Let's see what we're working with
```{r}
recipes %>% group_by(Type) %>% summarise(n_distinct(`Recipe ID`))
nofrosting %>% group_by(Type) %>% summarise(n_distinct(`Recipe ID`))
```
173 cupcake recipes and 101 muffin recipes.  If I exclude cupcake recipes with ambiguous frosting/cake designation for ingredients there are only 61 cupcake recipes!

# Calculate ammount/serving
```{r}
recipes <- recipes %>% mutate(amount_per = amount_sum/servings)
recipes %>% arrange((amount_per))
```


# Scatter plots
Great for spotting outliers
```{r}
cols <- c("cupcake" = alpha("blue", 0.5), "muffin" = alpha("red", 0.5))

p <- ggplot(recipes, aes(x = ingredient, y = amount_per, color = Type, label = `Recipe ID`)) +
  geom_point(alpha = 0.2) +
  coord_flip() +
  scale_color_manual(values = cols) +
  theme_bw()
p.1 <- gglabeller(p)
```
```{r}
p.1$plot
recipes[p.1$ix, ]
# recipes %>% filter(`Recipe ID` == "6869")
```

- **213813** is not a cupcake---it's "Steamed Pandan Tapioca Pearl Cakes"
- **218681** is a mini cheesecake---also not a cupcake
- 219859 is a legit cupcake, it just has a lot of artificial sweeteners of different types in it
- **24971** is "Chocolate Muck Muck Cake".  It has 4 eggs and 4 yolks in it.  I'm comfortable eliminating this one
- 260774 is a cupcake with a lot of weird toppings for ingredients.  It's legit.
- **6869** has duplicated ingredients for some reson.  Omit.
- **7164** call for "whole bran cereal" not just bran.  Omit.

```{r}
outliers <- c("6869", "24971", "218681", "213813", "7164")
recipes <- recipes %>% filter(!(`Recipe ID` %in% outliers))
```

# Ridge plots
```{r}
ggplot(recipes) +
  geom_density_ridges2(aes(x = amount_per, y = ingredient, fill = Type, color = Type),
                       alpha = 0.5,
                       rel_min_height = 0.005,
                       scale = 1.1) +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols) +
  #coord_cartesian(xlim = c(0,0.65)) + # This cuts off some recipes like one with 4 eggs.
  theme_bw()
```
# Heatmap
I think a heatmap might be more appropriate.  Gotta learn how to make one first
```{r}
recipes <- recipes %>% ungroup()
recipes_wide <-
  recipes %>%
  select(Type, `Recipe ID`, ingredient, amount_per) %>%
  mutate(ingredient = fct_infreq(ingredient)) %>% 
  spread(key = ingredient, value = amount_per)
recipes_wide
```
```{r}
recipes_matrix <- recipes_wide %>% select_if(is.double) %>% as.matrix()
rownames(recipes_matrix) <- recipes_wide$`Recipe ID`
```
```{r}
main_heatmap(recipes_matrix %>% scale(., center=FALSE, scale = TRUE),
             name = "amount per serving",
             x_categorical = TRUE,
             layout = list(font = list(size = 8))) %>%
  add_row_groups(recipes_wide$Type,
                  side = "left", name = "Muffin or Cupcake?",
                  title = "Muffin/<br>Cupcake",
                  colors = c(alpha("blue", 1), alpha("red", 1))) %>%
  add_col_labels()
```

231057 has a lot of vanilla, but legit
236467 has a lot of spice, but legit

Muffins seem to have less vanilla, less sugar, less cream cheese, and less "other"


# Calories
Are muffins or cupcakes more caloric on average?

```{r}
recipes %>% group_by(Type) %>% summarise(mean_cals = mean(calories, na.rm = TRUE))
```
```{r}
ggplot(recipes, aes(x = Type, y = calories)) +
  geom_boxplot() + theme_bw()
```

Woah! There is a cupcake with more than 800 calories per serving?!
```{r}
recipes %>% filter(calories > 750)
```
What's the most caloric muffin?
```{r}
recipes %>% filter(Type == "muffin") %>% summarise(ID = first(`Recipe ID`), calories = max(calories))
```

# Number of ingredients
How does number of ingredients vary by muffin/cupcake?

```{r}
recipes %>%
  group_by(Type, `Recipe ID`) %>%
  summarise(num_ingreds = n_distinct(ingredient)) %>% 
  group_by(Type) %>% 
  summarise(mean_ing_num = mean(num_ingreds))
```
Cupcakes have slightly more ingredients on average, probably because of frosting/toppings.  Oh, but this is after all the "others" are combined, so this might not be that meaningful.

# Make wide data for multivariate analysis
I need to calculate amount per serving for each ingredient and then `spread()` into a wide data frame
```{r}
recipes_wide <- recipes %>%
  mutate(amount_per_serving = amount_sum/servings) %>% 
  select(Type, `Recipe ID`, ingredient, amount_per_serving) %>% 
  spread(key = ingredient, value = amount_per_serving)
recipes_wide
```

This is a very sparse dataset with a LOT of `NA`s.  Let's see if I can figure out what percentage of the data is `NA`
```{r}
na_num <- recipes_wide %>% ungroup() %>% select_if(is.numeric) %>% is.na() %>% sum()
dimension <- recipes_wide %>% ungroup() %>% select_if(is.numeric) %>% dim()
na_num / (dimension[1] * dimension[2]) * 100
```
74% NA's.  Yikes!
```{r}
recipes_wide <- recipes_wide %>% 
  mutate_if(is.double, ~replace_na(.,0))
test <- prcomp(recipes_wide %>% select_if(is.double), scale. = TRUE)
summary(test)
biplot(test)
```

Needs 19 PCs to explain ~ 70% of variance.  Might be good for demonstrating benefits of OPLS vs PCA.

# Write outlier-removed recipes to RDS
```{r}
write_rds(recipes_wide, "recipes_wide.rds")
```

