---
title: "Cupcake V. Muffin EDA"
output: html_notebook
---

Now that I have a dataset of the ingredients of many cupcake and muffin recipes, it's time for phase 2---EDA.

I already know it is a pretty sparse data set and that the number of cupcakes and muffins are not very even.

Figures I might want to make:

- "heatmap" of pres/abs of ingredients.
- boxplot of calories with scatterplot and labeled outliers
- ridge plot of distribution of amounts per serving of each ingredient


**IMPORTANT** Don't use calories with cupcake recipes because you've removed the frosting!
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggridges)
```

# Read in files
```{r}
# With frosting and calories/serving
recipes <- read_rds("recipes_tidy.rds")
# without frosting and calories
nofrosting <- read_rds("nofrosting_tidy.rds")
```

# Number of Recipes
Let's see what we're working with
```{r}
recipes %>% group_by(Type) %>% summarise(n_distinct(`Recipe ID`))
nofrosting %>% group_by(Type) %>% summarise(n_distinct(`Recipe ID`))
```
173 cupcake recipes and 101 muffin recipes.  If I exclude cupcake recipes with ambiguous frosting/cake designation for ingredients there are only 61 cupcake recipes!


# Ridge plots
```{r}
recipes <- recipes %>% mutate(amount_per = amount_sum/servings)
recipes %>% arrange(desc(amount_per))

cols <- c("cupcake" = alpha("blue", 0.5), "muffin" = alpha("red", 0.5))

ggplot(recipes) +
  geom_density_ridges2(aes(x = amount_per, y = ingredient, fill = Type, color = Type),
                       #alpha = 0.5,
                       rel_min_height = 0.005,
                       scale = 1.1) +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols) +
  coord_cartesian(xlim = c(0,0.65)) + # This cuts off some recipes like one with 4 eggs.
  theme_bw()
```
# Heatmap
I think a heatmap might be more appropriate.  Gotta learn how to make one first



# Calories
Are muffins or cupcakes more caloric on average?

```{r}
recipes %>% group_by(Type) %>% summarise(mean_cals = mean(calories, na.rm = TRUE))
```
```{r}
ggplot(recipes, aes(x = Type, y = calories)) +
  geom_boxplot()
```

Woah! There is a cupcake with more than 800 calories per serving?!
```{r}
recipes %>% filter(calories > 750)
```
What's the most caloric muffin?
```{r}
recipes %>% filter(Type == "muffin") %>% ungroup() %>% summarise(ID = first(`Recipe ID`), cals = max(calories))
```

# Number of ingredients
How does number of ingredients vary by muffin/cupcake?

```{r}
recipes %>%
  group_by(Type, `Recipe ID`) %>%
  summarise(num_ingreds = n_distinct(ingredient)) %>% 
  group_by(Type) %>% 
  summarise(mean_ing_num = mean(num_ingreds))
```
Cupcakes have slightly more ingredients on average, probably because of frosting/toppings.  Oh, but this is after all the "others" are combined, so this might not be that meaningful.

# Make wide data for multivariate analysis
I need to calculate amount per serving for each ingredient and then `spread()` into a wide data frame
```{r}
recipes_wide <- recipes %>%
  mutate(amount_per_serving = amount_sum/servings) %>% 
  select(Type, `Recipe ID`, ingredient, amount_per_serving) %>% 
  spread(key = ingredient, value = amount_per_serving)
recipes_wide
```

This is a very sparse dataset with a LOT of `NA`s.  Let's see if I can figure out what percentage of the data is `NA`
```{r}
na_num <- recipes_wide %>% ungroup() %>% select_if(is.numeric) %>% is.na() %>% sum()
dimension <- recipes_wide %>% ungroup() %>% select_if(is.numeric) %>% dim()
na_num / (dimension[1] * dimension[2]) * 100
```
73% NA's.  Yikes!


