---
title: "Cupcake V. Muffin EDA"
output: html_notebook
---

Now that I have a dataset of the ingredients of many cupcake and muffin recipes, it's time for phase 2---EDA.

I already know it is a pretty sparse data set and that the number of cupcakes and muffins are not very even.

Figures I might want to make:

- "heatmap" of pres/abs of ingredients.
- boxplot of calories with scatterplot and labeled outliers
- ridge plot of distribution of amounts per serving of each ingredient


**IMPORTANT** Don't use calories with cupcake recipes because you've removed the frosting!

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggridges)
library(iheatmapr)
library(gglabeller)
library(here)
```

# Read in files
```{r}
# With frosting and calories/serving
recipes <- read_rds(here("data", "recipes_tidy.rds"))
# without frosting and calories
nofrosting <- read_rds(here("data", "nofrosting_tidy.rds"))
```

# Number of Recipes
Let's see what we're working with

```{r}
recipes %>% group_by(type) %>% summarise(n_distinct(recipe_id))
nofrosting %>% group_by(type) %>% summarise(n_distinct(recipe_id))
```
173 cupcake recipes and 101 muffin recipes.  If I exclude cupcake recipes with ambiguous frosting/cake designation for ingredients there are only 61 cupcake recipes!



# Scatter plots

Great for spotting outliers

```{r}
cols <- c("cupcake" = alpha("blue", 0.5), "muffin" = alpha("red", 0.5))

p <- ggplot(recipes, aes(x = ingredient, y = amount_per, color = type, label = recipe_id)) +
  geom_point(alpha = 0.2) +
  coord_flip() +
  scale_color_manual(values = cols) +
  theme_bw()
p.1 <- gglabeller(p, aes(label = recipes$recipe_id))
```

```{r}
p.1$plot
recipes[p.1$ix, ]
# recipes %>% filter(recipe_id == "242474")
```

- **213813** is not a cupcake---it's "Steamed Pandan Tapioca Pearl Cakes"
- **218681** is a mini cheesecake---also not a cupcake
- 219859 is a legit cupcake, it just has a lot of artificial sweeteners of different types in it
- 242474 appears to be a legit cupcake, but made with coconut flour.
- **24971** is "Chocolate Muck Muck Cake".  It has 4 eggs and 4 yolks in it.  I'm comfortable eliminating this one
- 255003 and 260789 are legit cupcakes.
- 79313 is a legit disgusting carrot cake cupcake that is mostly carrot.
- 260774 is a cupcake with a lot of weird toppings for ingredients.  It's legit.
- **6869** has duplicated ingredients for some reson.  Omit.
- **7164** call for "whole bran cereal" not just bran.  Omit.

```{r}
outliers <- c("6869", "24971", "218681", "213813", "7164")
recipes <- recipes %>% filter(!(recipe_id %in% outliers))
```

# Ridge plots

```{r}
ggplot(recipes) +
  geom_density_ridges2(aes(x = amount_per, y = ingredient, fill = type, color = type),
                       alpha = 0.5,
                       rel_min_height = 0.005,
                       scale = 1.1) +
  scale_fill_manual(values = cols) +
  scale_color_manual(values = cols) +
  theme_bw()
```

# Heatmap

I think a heatmap might be more appropriate.  Gotta learn how to make one first

```{r}
recipes <- recipes %>% ungroup()
recipes_wide <-
  recipes %>%
  select(type, recipe_id, ingredient, amount_per) %>%
  mutate(ingredient = fct_infreq(ingredient)) %>% 
  spread(key = ingredient, value = amount_per)
recipes_wide
```
```{r}
recipes_matrix <- recipes_wide %>% select_if(is.double) %>% as.matrix()
rownames(recipes_matrix) <- recipes_wide$recipe_id
```

```{r}
main_heatmap(recipes_matrix %>% scale(., center=FALSE, scale = TRUE),
             name = "amount per serving",
             x_categorical = TRUE,
             layout = list(font = list(size = 8))) %>%
  add_row_groups(recipes_wide$type,
                  side = "left", name = "Muffin or Cupcake?",
                  title = "Muffin/<br>Cupcake",
                  colors = c(alpha("blue", 1), alpha("red", 1))) %>%
  add_col_labels()
```

231057 has a lot of vanilla, but legit
236467 has a lot of spice, but legit

Muffins seem to have less vanilla, less sugar, less cream cheese, and less "other"


# Calories
Are muffins or cupcakes more caloric on average?

```{r}
recipes %>% group_by(type) %>% summarise(mean_cals = mean(calories, na.rm = TRUE))
```

```{r}
ggplot(recipes, aes(x = type, y = calories)) +
  geom_boxplot() + theme_bw()
```

Woah! There is a cupcake with more than 800 calories per serving?!

```{r}
recipes %>% filter(calories > 750)
```

What's the most caloric muffin?

```{r}
recipes %>% filter(type == "muffin") %>% summarise(ID = first(recipe_id), calories = max(calories))
```

# Number of ingredients

How does number of ingredients vary by muffin/cupcake?

```{r}
recipes %>%
  group_by(type, recipe_id) %>%
  summarise(num_ingreds = n_distinct(ingredient)) %>% 
  group_by(type) %>% 
  summarise(mean_ing_num = mean(num_ingreds))
```
Cupcakes have slightly more ingredients on average, probably because of frosting/toppings.  Oh, but this is after all the "others" are combined, so this might not be that meaningful.

# Make wide data for multivariate analysis

I need to calculate amount per serving for each ingredient and then `spread()` into a wide data frame

```{r}
recipes_wide <-
  recipes %>%
  select(type, recipe_id, calories_per, ingredient, amount_per) %>% 
  spread(key = ingredient, value = amount_per) %>% 
  ungroup()
recipes_wide

nofrosting_wide <- 
  nofrosting %>% 
  select(type, recipe_id, ingredient, amount_per) %>% 
  spread(key = ingredient, value = amount_per) %>% 
  ungroup()
nofrosting_wide
```

This is a very sparse dataset with a LOT of `NA`s.  Let's see if I can figure out what percentage of the data is `NA`

```{r}
na_num <- recipes_wide %>% ungroup() %>% select_if(is.numeric) %>% is.na() %>% sum()
dimension <- recipes_wide %>% ungroup() %>% select_if(is.numeric) %>% dim()
na_num / (dimension[1] * dimension[2]) * 100
```

74% NA's.  Yikes!


Replaces NAs with zeroes.  They're not truly NAs, that ingredient doesn't exist in the recipe if there is an NA.  EXCEPT for when it's calories.  There an NA really is an NA, but I'll just remove the one recipe with no info on calories.

```{r}
recipes_wide <- 
  recipes_wide %>% 
  filter(!is.na(calories_per)) %>% 
  mutate_if(is.double, ~replace_na(.,0))

nofrosting_wide <- 
  nofrosting_wide %>% 
  mutate_if(is.double, ~replace_na(.,0))
```

Needs 19 PCs to explain ~ 70% of variance.  Might be good for demonstrating benefits of OPLS vs PCA.

# Write outlier-removed recipes to RDS

```{r}
write_rds(recipes_wide, here("data", "recipes_wide.rds"))
write_rds(nofrosting_wide, here("data", "nofrosting_wide.rds"))
```

